From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/discussion.tex
Change sigma notation

\begin{center}
    \begin{tabular}{c c c}
        \hline
        $\PiLang$     & $\Sn$               & $\UFin$     \\
        \hline
        types         & Natural numbers     & Finite sets \\
        \hline
        1-combinators & Generators of $\Sn$ & 1-paths     \\
        \hline
        2-combinators & Relations of $\Sn$  & 2-paths     \\
        \hline
    \end{tabular}
\end{center}

\begin{center}
    \begin{tabular}{c c c}
        \hline
        Logic                              & Program                       & Algebra                     \\
        \hline
        Propositions                       & Types                         & Objects                     \\
        \hline
        Proofs                             & Terms                         & Morphisms                   \\
        \hline
        $\blank$                           & Equations                     & Homotopies                  \\
        \hline
        Intuitionistic propositional logic & Simply typed Lambda Calculus  & Cartesian Closed Categories \\
        \hline
        Superstructural Reversible logic   & Reversible Language $\PiLang$ & Symmetric Rig Groupoids     \\
        \hline
        Connectives                        & Type constructors             & Bifunctors                  \\
        \hline
        ?                                  & Type normalisation            & ?                           \\
        \hline
        ?                                  & Program normalisation         & ?                           \\
        \hline
        ?                                  & Soundness of 2-combinators    & ?                           \\
    \end{tabular}
\end{center}

\begin{center}
    \begin{tabular}{c c c}
        \hline
        Reversible Logic             & Reversible Programming                  & Algebra                  \\
        \hline
        Free commutative rig         & Free symmetric rig groupoid             & $\UFin$                  \\
        \hline
        Propositions                 & Types                                   & Objects                  \\
        \hline
        Entailment?                  & Reversible programs                     & Isomorphisms             \\
        \hline
        $\blank$                     & 2-combinators                           & Homotopies               \\
        \hline
        Conjunction?                 & $\times$                                & $\otimes$                \\
        \hline
        Disjunction?                 & $+$                                     & $\oplus$                 \\
        \hline
        $A \land B \vdash B \land A$ & $\swapt$                                & $\mathcal{B}_{A,B}$      \\
        \hline
        True proposition             & $\onet$                                 & $F_1$                    \\
        \hline
        Decidability                 & Program Synthesis                       & Reachability in a graph? \\
        \hline
        Soundness/Completeness       & Soundness/Completeness of 1-combinators & ?                        \\
    \end{tabular}
\end{center}

>>>>

% \begin{center}
%     \begin{tabular}{c c c}
%         \hline
%         $\PiLang$     & $\Sn$               & $\UFin$     \\
%         \hline
%         types         & Natural numbers     & Finite sets \\
%         \hline
%         1-combinators & Generators of $\Sn$ & 1-paths     \\
%         \hline
%         2-combinators & Relations of $\Sn$  & 2-paths     \\
%         \hline
%     \end{tabular}
% \end{center}

% \begin{center}
%     \begin{tabular}{c c c}
%         \hline
%         Logic                              & Program                       & Algebra                     \\
%         \hline
%         Propositions                       & Types                         & Objects                     \\
%         \hline
%         Proofs                             & Terms                         & Morphisms                   \\
%         \hline
%         $\blank$                           & Equations                     & Homotopies                  \\
%         \hline
%         Intuitionistic propositional logic & Simply typed Lambda Calculus  & Cartesian Closed Categories \\
%         \hline
%         Superstructural Reversible logic   & Reversible Language $\PiLang$ & Symmetric Rig Groupoids     \\
%         \hline
%         Connectives                        & Type constructors             & Bifunctors                  \\
%         \hline
%         ?                                  & Type normalisation            & ?                           \\
%         \hline
%         ?                                  & Program normalisation         & ?                           \\
%         \hline
%         ?                                  & Soundness of 2-combinators    & ?                           \\
%     \end{tabular}
% \end{center}

% \begin{center}
%     \begin{tabular}{c c c}
%         \hline
%         Reversible Logic             & Reversible Programming                  & Algebra                  \\
%         \hline
%         Free commutative rig         & Free symmetric rig groupoid             & $\UFin$                  \\
%         \hline
%         Propositions                 & Types                                   & Objects                  \\
%         \hline
%         Entailment?                  & Reversible programs                     & Isomorphisms             \\
%         \hline
%         $\blank$                     & 2-combinators                           & Homotopies               \\
%         \hline
%         Conjunction?                 & $\times$                                & $\otimes$                \\
%         \hline
%         Disjunction?                 & $+$                                     & $\oplus$                 \\
%         \hline
%         $A \land B \vdash B \land A$ & $\swapt$                                & $\mathcal{B}_{A,B}$      \\
%         \hline
%         True proposition             & $\onet$                                 & $F_1$                    \\
%         \hline
%         Decidability                 & Program Synthesis                       & Reachability in a graph? \\
%         \hline
%         Soundness/Completeness       & Soundness/Completeness of 1-combinators & ?                        \\
%     \end{tabular}
% \end{center}

### 2021-07-09 Pi+/paper/ufin.tex
Fixes
Change sigma notation

\begin{propositionrep}[Univalent Subuniverse]
  \label{prop:univsub}
  A universe predicate is a type family $P : \UU \to \UU$ whose fibers are propositions, that is, $P(X)$ is a
  proposition for every $X$. Given such a predicate, the fibration ${\fst : \dsum*{X:\UU}{P(X)} \to \UU}$ is univalent
  and generates a univalent subuniverse $\UU_{P} \defeq (\dsum{X:\UU}{P(X)}, \fst)$.~\footnote{Univalent
    typoids~\cite{petrakisUnivalentTypoids2019a} are a different presentation of univalent subuniverses.}
\end{propositionrep}

@@

\begin{definition}[$\Fin$]
  \label{def:finite-set}
  The type family $\Fin : \Nat \to \UU$ is the type of finite sets indexed by their cardinality. It is defined
  equivalently in two different ways,
  \begin{gather*}
    \Fin[n] \defeq \dsum{k:\Nat}{k < n}
    \qquad\qquad \text{or} \qquad\qquad
    \begin{aligned}
       & \Fin[0] \defeq \bot                      \\
       & \Fin[\suc[n]] \defeq \top \sqcup \Fin[n]
    \end{aligned}
  \end{gather*}
  Note that $\Fin[n]$ is a set, and we use both definitions interchangeably.
\end{definition}

\begin{definition}[$\isFin$]
  \label{def:isfin}
  We say that a type is finite if it is merely equal to $\Fin[n]$ for some $n : \Nat$.
  \[
    \isFin[X] \defeq \dsum{n:\Nat}{\SubP{X}{\Fin[n]}}
  \]
  Note that the natural number $n$ need not be truncated, as justified below.
\end{definition}

@@

\begin{definition}
  \label{def:ufin}
  The univalent subuniverse of \emph{all finite types} is given by
  $
    \UFin \defeq \dsum{X:\UU}{\isFin[X]}.
  $
  We write $F_{n} \defeq (\Fin[n], n, \trunc{\refl})$, for the image of the inclusion of $\Fin[n]$.
\end{definition}

@@

\begin{definition}[$\UFin[n]$]
  For any $n : \Nat$, we define $\UFin[n] \defeq \BAut[\Fin[n]]$ to be the univalent subuniverse of $n$-element sets.
  Note that, $\UFin$ can be equivalently seen as the collection of all types of finite cardinality, that is,
  $\UFin \eqv \dsum*{n:\Nat}{\UFin[n]}$.
\end{definition}

>>>>

\begin{propositionrep}[Univalent Subuniverse]
  \label{prop:univsub}
  A universe predicate is a type family $P : \UU \to \UU$ whose fibers are propositions, that is, $P(X)$ is a
  proposition for every $X$. Given such a predicate $P$, the fibration ${\fst : \dsum*{X:\UU}{P(X)} \to \UU}$ is
  univalent and generates a univalent subuniverse ${\UU_{P} \defeq (\dsum*{X:\UU}{P(X)}, \fst)}$.~\footnote{Univalent
    typoids~\cite{petrakisUnivalentTypoids2019a} are a different presentation of univalent subuniverses.}
\end{propositionrep}

@@

\begin{definition}[$\Fin$]
  \label{def:finite-set}
  The type family $\Fin : \Nat \to \UU$ is the type of finite sets indexed by their cardinality. It is defined
  equivalently in two different ways,
  \begin{gather*}
    \Fin[n] \defeq \dsum*{k:\Nat}{k < n}
    \qquad\qquad \text{or} \qquad\qquad
    \begin{aligned}
       & \Fin[0] \defeq \bot                      \\
       & \Fin[\suc[n]] \defeq \top \sqcup \Fin[n]
    \end{aligned}
  \end{gather*}
  Note that $\Fin[n]$ is a set, and we use both definitions interchangeably.
\end{definition}

\begin{definition}[$\isFin$]
  \label{def:isfin}
  We say that a type is finite if it is merely equal to $\Fin[n]$ for some $n : \Nat$.
  \[
    \isFin[X] \defeq \dsum*{n:\Nat}{\SubP{X}{\Fin[n]}}
  \]
  Note that the natural number $n$ need not be truncated, as justified below.
\end{definition}

@@

\begin{definition}
  \label{def:ufin}
  The univalent subuniverse of \emph{all finite types} is given by
  $
    \UFin \defeq \dsum*{X:\UU}{\isFin[X]}.
  $
  We write $F_{n} \defeq (\Fin[n], n, \trunc{\refl})$, for the image of the inclusion of $\Fin[n]$.
\end{definition}

@@

\begin{definition}[${\UFin[n]}$]
  For any $n : \Nat$, we define $\UFin[n] \defeq \BAut[\Fin[n]]$ to be the univalent subuniverse of $n$-element sets.
  Note that, $\UFin$ can be equivalently seen as the collection of all types of finite cardinality, that is,
  $\UFin \eqv \dsum*{n:\Nat}{\UFin[n]}$.
\end{definition}
